---
title: "Analysis of Project Runover Data"
---
```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
hss <- read.csv("Project_RunOver.csv") %>% na.omit()
temp <- str_split(hss$Date.of.Encounter, "/", simplify = T)
temp[,3] <- ifelse(str_length(temp[,3]) == 2, paste0("20", temp[,3]), temp[,3])
hss$Date.of.Encounter <- ymd(paste0(temp[,3], "-", temp[,1], "-", temp[,2]))
hss$year_val <- year(hss$Date.of.Encounter)

df <- group_by(hss %>% count(Which.of.the.following.groups.best.describes.the.roadkill.found.))
```

# Temporal Analysis

#### To see how roadkill numbers have changed over the years, take a look at this line plot: 

```{r, echo=FALSE}
yearly_counts <- hss %>% na.omit("hss") %>% filter(year_val != 2024) %>%
  group_by(year_val) %>%
  summarise(Count = n())

# Plotting the data
ggplot(yearly_counts, aes(x = year_val, y = Count)) +
  geom_line(group=1, color="blue") +
  geom_point(color="red") +
  scale_x_continuous(breaks = yearly_counts$year_val, labels = yearly_counts$year_val) + # Set x-axis breaks and labels
  labs(title="Yearly Roadkill Incidents", x="Year", y="Count") +
  theme_minimal()

```

<p> As seen in the graph, the number of reptile roadkill has increased exponentially from 2011 to 2023. This shows us that wildlife roadkill is a legitimate and pressing concern. </p>

<p> The low records from 2011 to 2020 is because HSS had not officially launched Project Runover yet. Project Runover was only launched in 2020 which explains the sharp increase from 2020 to 2021. </p>

<p> However, as the project gained traction in 2021, the number of yearly roadkill accidents has steadily increased, emphasising the need for mitigation measures to prevent numbers from climbing. </p>

<p> For this graph, I did not include the year 2024 as data for this year is still being collected and accurate data is not available yet. </p>

# Species Analysis

#### Let's zoom in to see which types of reptiles are the most affected by roadkill:

```{r echo=FALSE, warning=FALSE}
plot <- ggplot(df, aes(x = 
Which.of.the.following.groups.best.describes.the.roadkill.found., y = n, fill = Which.of.the.following.groups.best.describes.the.roadkill.found.)) + geom_bar(stat = "identity") + labs(title = "Number of Each Reptile Roadkill by Reptile Type ", x = "Reptile Type", y = "Number of Reptile Roadkill") + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + scale_fill_viridis_d() +
  guides(fill = "none")
plot
```

<p> As seen in the bar plot, snakes are the most affected by roadkill, followed by frogs and monitor lizards. Reptiles are vulnerable to vehicle collisions because they often crawl onto roads to regulate their body temperature. The high occurrence of snake roadkill could be due to stigma against snakes, motivating drivers to run over them. </p>

#### Interested to know how roadkill incidents have affected the different reptiles each year? Choose a reptile you're interested to learn more about!

```{=html}
<iframe height="500" width="100%" frameborder="no" src="https://ruiwenlim.shinyapps.io/MyShinyApp2/"> </iframe>
```

#### Looking at the time of year may also paint a picture on the factors affecting reptile roadkill. 

```{r, echo=FALSE, warning=FALSE}
hss %>%
  mutate(month = month(Date.of.Encounter)) %>%
  group_by(Which.of.the.following.groups.best.describes.the.roadkill.found., month) %>%
  count() %>%
  ggplot(aes(x = month, y = Which.of.the.following.groups.best.describes.the.roadkill.found., fill = n)) +
  geom_tile() +
  labs(x = "Month", y = "Species", fill = "Count") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 25) +
  theme_bw() +
  scale_x_continuous(breaks = 1:12, labels = month.abb)
```

<p> As seen from the heat map, the month with the highest snake roadkill was from May to June. This period is usually the hottest month in Singapore. As such, there may be increased instances of snakes seeking shade beyond forested areas. Other factors such as mating season may affect the number of reptile roadkill found that month. </p>

#### Curious about how some of these reptiles look? Take a look at the first 5 reptiles recorded in Project Runover:

```{=html}
<iframe height="500" width="100%" frameborder="no" src="https://ruiwenlim.shinyapps.io/snakeapp/"> </iframe>
```

# Spatial Analysis

#### This interactive map allows you to see the locations where reptile roadkill have been spotted. Using the slider on the left, you can select the year(s). The blue circles on the map represent the location of the reptile roadkill. Can you spot any reptile roadkill hotspots?

```{=html}
<iframe height="500" width="100%" frameborder="no" src="https://ruiwenlim.shinyapps.io/MyShinyApp/"> </iframe>
```

<p> Ironically, but not surprisingly, some reptile roadkill hotspots are roads near nature reserves and parks. Nature reserves and sparks are rich habitats for biodiversity. Due to the disruption of forest connectivity to build these roads, wildlife often use roads to move from one place to another. The high concentration of wildlife in these nature reserves also means an increased probability of roadkill. When driving in these areas, drive slowly and keep a lookout for any wildlife! </p>

#### Now that we know some reptile roadkill hotspots, we can see what type of paths these incidents usually occur on.  

```{r, echo=FALSE}
# Aggregate data by type of path
path_counts <- hss %>%
  group_by(How.would.you.classify.the.path.this.roadkill.was.found.on.) %>%
  summarise(Count = n()) %>%
  arrange(desc(Count))  # Optional: arrange in descending order of count
# Creating a bar chart
path_counts %>%
  slice_max(order_by = Count, n = 8) %>%
  filter(How.would.you.classify.the.path.this.roadkill.was.found.on. != "") %>%
  mutate(How.would.you.classify.the.path.this.roadkill.was.found.on. = reorder(How.would.you.classify.the.path.this.roadkill.was.found.on., -Count)) %>%
  ggplot(aes(x = How.would.you.classify.the.path.this.roadkill.was.found.on., y = Count, fill = How.would.you.classify.the.path.this.roadkill.was.found.on.)) +
  geom_bar(stat = "identity", color = "black") +  # Use identity to use Count values as heights
  labs(title = "Roadkill Incidents by Type of Path",
       x = "Type of Path",
       y = "Number of Incidents") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x labels for better
  guides(fill = "none") +
  scale_fill_viridis_d()

```

#### As seen by the graph, a very large majority of reptile roadkill are found on roads. This emphasises the adverse impacts urbanisation, and more specifically, roads have had on wildlife. 

<p> Through exploring this data, it is reasonable to conclude that urbanisation has indeed had a devastating impact on wildlife. However, these roadkill numbers don't have to be this high! To take action, you can start by contributing to Project Runover if you see any roadkill, driving slowly and staying alert especially on roads near nature reserves, and feedback to governmental institutions about mitigation measures! </p>

<p><strong> TOGETHER, WE CAN STOP REPTILE ROADKILL! </strong></p>

#### References
<div style="text-indent: -0.5in; padding-left: 0.5in;">

[Banded Malayan Coral Snake—Calliophis intestinalis. (n.d.).](https://www.ecologyasia.com/verts/snakes/banded_malayan_coral_snake.htm) Retrieved 15 April 2024.

[Common wolf snake (Lycodon capucinus). (n.d.).](https://www.thainationalparks.com/species/lycodon-capucinus) Retrieved 15 April 2024.

[Dendrelaphis cyanochloris. (2022, September 19). In Wikipedia.](https://en.wikipedia.org/wiki/Dendrelaphis_cyanochloris)

[Indochinese Rat Snake—Ptyas korros. (n.d.).](https://www.ecologyasia.com/verts/snakes/indo-chinese_rat_snake.htm) Retrieved 15 April 2024.

[Oriental Whipsnake (Snakes of the Philippines) · iNaturalist. (n.d.). iNaturalist.](https://www.inaturalist.org/guide_taxa/1180819) Retrieved 15 April 2024.

</div>









